---
title: "GWAS results"
author: "Yunqi Yang"
date: "6/13/2024"
output: html_document
---

## Description:

1. Logistic gwas result for all asthma, childhood asthma and adulthood asthma

2. Survival gwas result for all asthma, childhood asthma and adulthood asthma

3. Compare logistic and survival gwas on all asthma

4. Compare logistic and survival gwas on COA

5. Compare logistic and survival gwas on AOA

```{r}
library(stringr)
```

### 1. Logistic gwas
```{r}
file.aoa <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/aoa", pattern = "\\.glm\\.logistic$", full.names = TRUE)
file.coa <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/coa", pattern = "\\.glm\\.logistic$", full.names = TRUE)

file <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/all", pattern = "\\.glm\\.logistic$", full.names = TRUE)
```

```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))
for (i in 1:length(file)){
  region = file[i]
  region_name = strsplit(region, "/")[[1]][9]
  gwas = read.csv(region, header = TRUE, sep = "\t")
  plot(gwas$POS, -log10(gwas$P), pch = 20, cex = 0.5, main = region_name, ylab = "log10-pval", xlab = "Position")
}

```


```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))
for (i in 1:length(file.aoa)){
  region.aoa = file.aoa[i]
  region.coa = file.coa[i]
  region_name = strsplit(region.aoa, "/")[[1]][9]
  gwas.aoa = read.csv(region.aoa, header = TRUE, sep = "\t")
  gwas.coa = read.csv(region.coa, header = TRUE, sep = "\t")
  ymin = min(log10(gwas.aoa$P)) - 1
  ymax = max(-log10(gwas.coa$P)) + 2
  plot(gwas.aoa$POS, log10(gwas.aoa$P), ylim = c(ymin, ymax), pch = 20, cex = 0.5, col = "#2c7fb8", main = region_name, ylab = "log10-pval", xlab = "Position")
  points(gwas.coa$POS, -log10(gwas.coa$P), pch = 20, cex = 0.5, col = "#de2d26")
  legend("topright", legend = c("COA", "AOA"), col = c("#de2d26", "#2c7fb8"), pch = 20)
}

```








### 2. Survival gwas

```{r}
path_to_files = "/Users/nicholeyang/Downloads/survivalsusie/result/gwas_surv"
file.aoa <- list.files(path = path_to_files, pattern = "^aoa.*\\.rds$", full.names = TRUE)
file.coa <- list.files(path = path_to_files, pattern = "^coa.*\\.rds$", full.names = TRUE)
file <- list.files(path = path_to_files, pattern = "^all.*\\.rds$", full.names = TRUE)

```

```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))
pattern <- "chr[0-9XYM]+_[0-9]+_[0-9]+"
for (i in 1:length(file)){
  region = file[i]
  region_name = str_extract(region, pattern)
  gwas = readRDS(region)
  plot(-log10(gwas[, 'p.value.spa']), pch = 20, cex = 0.5, main = region_name, ylab = "log10-pval", xlab = "Position")
}

```


```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))
for (i in 1:length(file.aoa)){
  region.aoa = file.aoa[i]
  region.coa = file.coa[i]
  region_name = str_extract(region.aoa, pattern)
  gwas.aoa = readRDS(region.aoa)
  gwas.coa = readRDS(region.coa)
  ymin = min(log10(gwas.aoa[, 'p.value.spa']), na.rm = TRUE) - 1
  ymax = max(-log10(gwas.coa[, 'p.value.spa']), na.rm = TRUE) + 2
  plot(log10(gwas.aoa[, 'p.value.spa']), ylim = c(ymin, ymax), pch = 20, cex = 0.5, col = "#2c7fb8", main = region_name, ylab = "log10-pval", xlab = "Position")
  points(-log10(gwas.coa[, 'p.value.spa']), pch = 20, cex = 0.5, col = "#de2d26")
  legend("topright", legend = c("COA", "AOA"), col = c("#de2d26", "#2c7fb8"), pch = 20)
}

```

```{r fig.width=10, fig.height=8}
pdf("/Users/nicholeyang/Downloads/survival-data-analysis/output/Fig_gwas_coa_aoa.pdf", width = 10, height = 8)
par(mfrow = c(2,2))
indx = c(1, 2, 4, 9)
labels = c("(a)", "(b)", "(c)", "(d)")
region_name = c("1q21.3", "10p14", "12q13.11", "2q12.1")
for (i in 1:length(indx)){
  region.aoa = file.aoa[indx[i]]
  region.coa = file.coa[indx[i]]
  gwas.aoa = readRDS(region.aoa)
  gwas.coa = readRDS(region.coa)
  
  ymin = min(log10(gwas.aoa[, 'p.value.spa']), na.rm = TRUE) - 1
  ymax = max(-log10(gwas.coa[, 'p.value.spa']), na.rm = TRUE) + 2
  
  plot(log10(gwas.aoa[, 'p.value.spa']), ylim = c(ymin, ymax), pch = 20, cex = 0.5, col = "#2c7fb8", main = region_name[i], ylab = "-log10(P value)", xlab = "", axes = FALSE)
  points(-log10(gwas.coa[, 'p.value.spa']), pch = 20, cex = 0.5, col = "#de2d26")
  
  # Custom y-axis ticks and labels
  y_ticks_pos <- seq(0, ymax, by = 5)
  y_ticks_neg <- seq(0, abs(ymin), by = 5)
  
  axis(2, at = c(-rev(y_ticks_neg), y_ticks_pos), labels = c(rev(y_ticks_neg), y_ticks_pos), las = 1)
  
  box()  # Adding box around the plot
  
  legend("topright", legend = c("COA", "AOA"), col = c("#de2d26", "#2c7fb8"), pch = 20)
  mtext("Position", side = 1, line = 3, adj = 0.5, cex = 0.8)
  mtext(labels[i], side = 1, line = 4, adj = 0.5, cex = 1)
}

```







### 3. Compare logistic gwas and survival gwas 

```{r}
file.logistic <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/all", pattern = "\\.glm\\.logistic$", full.names = TRUE)
file.survival <- list.files(path = "/Users/nicholeyang/Downloads/survivalsusie/result/gwas_surv", pattern = "^all.*\\.rds$", full.names = TRUE)
```



```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))

pattern <- "chr[0-9XYM]+_[0-9]+_[0-9]+"
for (i in 1:length(file.survival)){
  region = file.survival[i]
  region_name = str_extract(region, pattern)

  path.logistic = paste0("/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/all/gwas_", 
                         region_name, ".asthma.glm.logistic")
  
  # read in gwas results
  gwas.logistic = read.csv(path.logistic, header = TRUE, sep = "\t")
  gwas.survival = data.frame(readRDS(region))
  
  # merge two datasets
  snp_ids = sapply(rownames(gwas.survival), function(x) unlist(strsplit(x, "_"))[1])
  gwas.survival$ID = snp_ids
  dat = merge(gwas.logistic, gwas.survival, by = "ID")
  plot(dat$POS, -log10(dat$P), pch = 20, cex = 0.5, main = region_name, ylab = "log10-pval", xlab = "Position", ylim = c(0, max(-log10(dat$P), -log10(dat$p.value.spa)) + 1), col = "#de2d26")
  points(dat$POS, -log10(dat$p.value.spa), pch = 20, cex = 0.5)
  
  legend("topright", legend = c("logistic.gwas", "survival.gwas"), col = c("#de2d26", 1), pch = 20)
}

```










### 4. Compare logistic gwas and survival gwas on COA

```{r}
file.logistic <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/coa", pattern = "\\.glm\\.logistic$", full.names = TRUE)
file.survival <- list.files(path = "/Users/nicholeyang/Downloads/survivalsusie/result/gwas_surv", pattern = "^coa.*\\.rds$", full.names = TRUE)
```



```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))

pattern <- "chr[0-9XYM]+_[0-9]+_[0-9]+"
for (i in 1:length(file.survival)){
  region = file.survival[i]
  region_name = str_extract(region, pattern)

  path.logistic = paste0("/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/coa/gwas_", 
                         region_name, ".asthma_coa.glm.logistic")
  
  # read in gwas results
  gwas.logistic = read.csv(path.logistic, header = TRUE, sep = "\t")
  gwas.survival = data.frame(readRDS(region))
  
  # merge two datasets
  snp_ids = sapply(rownames(gwas.survival), function(x) unlist(strsplit(x, "_"))[1])
  gwas.survival$ID = snp_ids
  dat = merge(gwas.logistic, gwas.survival, by = "ID")
  plot(dat$POS, -log10(dat$P), pch = 20, cex = 0.5, main = region_name, ylab = "log10-pval", xlab = "Position", ylim = c(0, max(-log10(dat$P), -log10(dat$p.value.spa), na.rm = TRUE) + 1), col = "#de2d26")
  points(dat$POS, -log10(dat$p.value.spa), pch = 20, cex = 0.5)
  
  legend("topright", legend = c("logistic.gwas", "survival.gwas"), col = c("#de2d26", 1), pch = 20)
}

```

### 5. Compare logistic gwas and survival gwas on AOA

```{r}
file.logistic <- list.files(path="/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/aoa", pattern = "\\.glm\\.logistic$", full.names = TRUE)
file.survival <- list.files(path = "/Users/nicholeyang/Downloads/survivalsusie/result/gwas_surv", pattern = "^aoa.*\\.rds$", full.names = TRUE)
```



```{r fig.width=10, fig.height=20}

par(mfrow = c(6,2))

pattern <- "chr[0-9XYM]+_[0-9]+_[0-9]+"
for (i in 1:length(file.survival)){
  region = file.survival[i]
  region_name = str_extract(region, pattern)

  path.logistic = paste0("/Users/nicholeyang/Downloads/survivalsusie/data/gwas_logistic_out/aoa/gwas_", 
                         region_name, ".asthma_aoa.glm.logistic")
  
  # read in gwas results
  gwas.logistic = read.csv(path.logistic, header = TRUE, sep = "\t")
  gwas.survival = data.frame(readRDS(region))
  
  # merge two datasets
  snp_ids = sapply(rownames(gwas.survival), function(x) unlist(strsplit(x, "_"))[1])
  gwas.survival$ID = snp_ids
  dat = merge(gwas.logistic, gwas.survival, by = "ID")
  plot(dat$POS, -log10(dat$P), pch = 20, cex = 0.5, main = region_name, ylab = "log10-pval", xlab = "Position", ylim = c(0, max(-log10(dat$P), -log10(dat$p.value.spa), na.rm = TRUE) + 1), col = "#de2d26")
  points(dat$POS, -log10(dat$p.value.spa), pch = 20, cex = 0.5)
  
  legend("topright", legend = c("logistic.gwas", "survival.gwas"), col = c("#de2d26", 1), pch = 20)
}

```

















