---
title: "Self-report asthma: Coxph-SuSIE"
author: "Yunqi Yang"
date: "6/20/2024"
output: html_document
---


## Description:

Coxph Susie result on all asthma/ AOA/ COA in UKBiobank. 


```{r}
library(survival)
library(susieR)
devtools::load_all("/Users/nicholeyang/Downloads/logisticsusie")
```

### Region 1

Strong signals for COA, marginal significant for AOA. rs61894547 was the most significant SNP reported by Carole's paper, but not the most significant one in my result. However, have the largest PIP. 

#### 1. All asthma cases
```{r}
region = "chr11_75500001_77400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 2")

```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```


#### 2. COA
```{r}
region = "chr11_75500001_77400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 2")

```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```

#### 3. AOA
```{r}
region = "chr11_75500001_77400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")

```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```
```{r}
rm()
```


### Region 2

No GWAS significant signal for COA, marginal significant for AOA. 

Result: for all asthma and COA, no CS found. For AOA, there is one CS. rs56389811_T was the top significant signal reported by Carole's paper, and also the top significant one found in AOA survival gwas. PIP = 0.2


#### 1. All asthma cases
```{r}
region = "chr12_46000001_48700000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

#### 2. COA
```{r}
region = "chr12_46000001_48700000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

#### 3. AOA
```{r}
region = "chr12_46000001_48700000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")

```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```


```{r}
rm(res, gwas, X, fit)
```

### Region 3

Very strong signals for COA, very week signals for AOA. 

#### 1. All asthma cases
```{r}
region = "chr17_33500001_39800000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 2")

```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```
```{r}
rm(res, gwas, X, fit)
```

#### 2. COA
```{r}
region = "chr17_33500001_39800000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```

```{r}
rm(res, gwas, X, fit)
```


#### 3. AOA
```{r}
region = "chr17_33500001_39800000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```
```{r}
rm(res, gwas, X, fit)
```









### Region 4

Marginal significant signals for both COA and AOA. Combined analysis a lot more significant. 

#### 1. All asthma cases
```{r}
region = "chr10_6600001_12200000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r fig.height = 10, fig.width=8}
par(mfrow = c(3,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 2")

snps3 = colnames(X)[cs$cs$L3]
colors <- ifelse(rownames(gwas) %in% snps3, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 3")

snps5 = colnames(X)[cs$cs$L5]
colors <- ifelse(rownames(gwas) %in% snps5, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 5")


snps6 = colnames(X)[cs$cs$L6]
colors <- ifelse(rownames(gwas) %in% snps6, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 6")

snps7 = colnames(X)[cs$cs$L7]
colors <- ifelse(rownames(gwas) %in% snps7, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 7")



```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
cbind(gwas[rownames(gwas) %in% snps3, ], pip[sort(cs$cs$L3)])
cbind(gwas[rownames(gwas) %in% snps5, ], pip[sort(cs$cs$L5)])
cbind(gwas[rownames(gwas) %in% snps6, ], pip[sort(cs$cs$L6)])
cbind(gwas[rownames(gwas) %in% snps7, ], pip[sort(cs$cs$L7)])
```


```{r}
rm(res, gwas, X, fit)
```

#### 2. COA
```{r}
region = "chr10_6600001_12200000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
snps3 = colnames(X)[cs$cs$L3]
colors <- ifelse(rownames(gwas) %in% snps3, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 3")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps3, ], pip[sort(cs$cs$L3)])
```

```{r}
rm(res, gwas, X, fit)
```


#### 3. AOA
```{r}
region = "chr10_6600001_12200000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
```


```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```

```{r}
rm(res, gwas, X, fit)
```








### Region 5

COA: pval = 1e-40, AOA no significant signals. 

#### 1. All asthma cases
```{r}
region = "chr1_150600001_155100000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r }
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 2")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```


```{r}
rm(res, gwas, X, fit)
```

#### 2. COA
```{r}
region = "chr1_150600001_155100000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 2")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```

```{r}
rm(res, gwas, X, fit)
```


#### 3. AOA
```{r}
region = "chr1_150600001_155100000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```
