---
title: "Self-report asthma: Coxph-SuSIE"
author: "Yunqi Yang"
date: "6/20/2024"
output: html_document
---


## Description:

Coxph Susie result on all asthma/ AOA/ COA in UKBiobank. 


```{r}
library(survival)
library(susieR)
devtools::load_all("/Users/nicholeyang/Downloads/logisticsusie")
```

### Region 1

Marginal significant signals for COA, weak signals for AOA. 

rs11071559_T was the one with smallest pvalue in all asthma, and PIP = 0.24. Carole's paper also reported this one as the top signal. But in AOA, it's not the one with smallest pval, the pip is a lot smaller. 

#### 1. All asthma cases
```{r}
region = "chr15_59000001_63400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```

```{r}
rm(res, gwas, X, fit)
```

#### 2. COA
```{r}
region = "chr15_59000001_63400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```


```{r}
rm(res, gwas, X, fit)
```

#### 3. AOA
```{r}
region = "chr15_59000001_63400000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
rm(res, gwas, X, fit)
```



### Region 2

Very significant signals for COA, marginal significant signals for AOA. 

All asthma has a very weird CS. One All asthma CS overlap with COA CS. AOA CS has no overlap with other CSs. 


#### 1. All asthma cases
```{r}
region = "chr2_102100001_105300000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/all/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/all_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]
print(res[[3]])

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
par(mfrow = c(1,2))
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")

snps2 = colnames(X)[cs$cs$L2]
colors <- ifelse(rownames(gwas) %in% snps2, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 2")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
cbind(gwas[rownames(gwas) %in% snps2, ], pip[sort(cs$cs$L2)])
```

```{r}
rm(res, gwas, X, fit)
```

#### 2. COA
```{r}
region = "chr2_102100001_105300000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/coa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/coa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.5, pch = 20, main = "CS 1")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```


```{r}
rm(res, gwas, X, fit)
```

#### 3. AOA
```{r}
region = "chr2_102100001_105300000"
res = readRDS(paste0("/Users/nicholeyang/Downloads/survivalsusie/result/asthma_self_report/result/aoa/fit.susie.", region, ".rds"))
gwas = readRDS(paste0("/Users/nicholeyang/downloads/survivalsusie/result/gwas_surv/aoa_gwas_", region, ".rds"))

```

```{r}
fit = res[[1]]
X = res[[2]]

pip <- logisticsusie:::get_pip(fit$alpha)
effect_estimate <- colSums(fit$alpha * fit$mu)

class(fit) = "susie"
cs <- susie_get_cs(fit, X)
cs
```

```{r}
snps1 = colnames(X)[cs$cs$L1]
colors <- ifelse(rownames(gwas) %in% snps1, "red", "black")
plot(-log10(gwas[, "p.value.spa"]), col = colors, xlab = "SNP", ylab = "-log10(p-value)", cex = 0.8, pch = 20, main = "CS 1")
```

```{r}
cbind(gwas[rownames(gwas) %in% snps1, ], pip[sort(cs$cs$L1)])
```


```{r}
rm(res, gwas, X, fit)
```
